---
title: "Reproducible workflow for the project"
subtitle: "\"Multi-modal immune phenotyping of maternal peripheral blood in normal human pregnancy\""
author: "Yuri Kotliarov, Richard Apps"
output: 
  html_notebook: 
    toc: yes
editor_options: 
  chunk_output_type: inline
---

# Introduction

The following workflow was used to analyze the data and generate the figures for our paper "Multi-modal immune phenotyping of maternal peripheral blood in normal human pregnancy" to be published in JCI Insight in 2020.

<br/>

The workflow requires several R packages. To install run these commands in R:

```
# CRAN:
install.packages(c("tidyverse", "data.table", "readxl", "circlize"), repos = "https://cran.rstudio.com")

# Bioconductor:
install.packages("BiocManager")
BiocManager::install()
BiocManager::install(c("fgsea", "corrplot", "ComplexHeatmap", "survcomp""))
install.packages("eNetXplorer")

# Somalogic:
devtools::install_github("cran/pathological")
BiocManager::install("readat") 

```

# Data directory

The data directory should contain the following files:

* **Pregnancy_clinical_data.xlsx** - Clinical data for study participants, general and measured at each visits, including CBC and TBNK clinical flow
* **Pregnancy_flow_cytometry.xlsx** - Flow cytometry data: gated cell population frequencies, annotation and selected populations for Figure 2
* **Pregnancy.hybNorm.plateScale.calibrate.20170512.adat** - Somalogic data in ADAT file after normalization for hybridization  batch, scaling by plates and calibration against reference data
* **Model_genes.xlsx** - model genes from Aghaeepour et al, American journal of obstetrics and gynecology (2018)
* **pred.soma.old.txt** - list of 8 proteins in reduced model published in  Aghaeepour et al (2018)
* **pred.soma.new.txt** - list of 7 proteins in the new model (provided for reproducibility)

<br/>

# Generated data and figures

The new directories will be created by the workflow:

  * **data_generated** - input data reformatted or prefiltered
  * **results** - results of the analyses
  * **figures** - all fiures in PNG or PDF foremat are stored here

# Plots visits map (Figure 1)

```{r}
source("R/plot_visits.r")
```

# Prepare clinical data from further analysis

```{r}
source("R/clinical_data_prep.r")
```


# Flow data analysis (Figure 2)

```{r}
source("R/flow_analysis.r")
```

# Somalogic data pre-processing and analysis

```{r}
source("R/soma_analysis.r")
```

# Correlating Somalogic and flow cytometry data (Supp. Fig. 1)

```{r}
source("R/soma_vs_flow.r")
```

# Prepare Somalogic data for elastic net analysis

```{r}
source("R/soma_enet_prep.r")
```

# Run eNetXplorer for previous model of 70 genes (Figure 3A)

```{r}
source("R/Soma_70genes_enet_run.r")
source("R/Soma_70genes_enet_FIGURE.r")
```

# Run eNetXplorer for previous reduced model of 8 genes (Figure 3B)

```{r}
source("R/Soma_8genes_enet_run.r")
source("R/Soma_8genes_enet_FIGURE.r")
```

# Run eNetXplorer for all genes and check the erichment of genes from previous models (Supp. Fig. 2)

```{r}
source("R/Soma_all.genes_enet_run.r")
source("R/Soma_all.genes_enet_enrichment.r")
```


# Run eNetXplorer removing 70 genes from previous model (Figure 3C,D)

```{r}
source("R/Soma_70genes.removed_enet_run.r")
source("R/Soma_70genes.removed_enet_FIGURE.r")
```

# Correlate data for 8 genes in the old model and 7 genes in the new model

```{r}
source("R/soma_8_vs_7_correlation.r")
```

# Run eNetXplorer for Clinical data including CBC and TBNK clinical flow (Figure 3F,G)

```{r}
source("R/Clinical_enet_run.r")
source("R/Clinical_enet_FIGURE.r")
```

# Run eNetXplorer for CBC data only (Supp. Fig. 3)

```{r}
source("R/CBC_enet_run.r")
source("R/CBC_enet_FIGURE.r")
```


---

```
> sessionInfo()
R version 3.6.1 (2019-07-05)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 17763)

Matrix products: default

locale:
[1] LC_COLLATE=English_United States.1252  LC_CTYPE=English_United States.1252    LC_MONETARY=English_United States.1252
[4] LC_NUMERIC=C                           LC_TIME=English_United States.1252    

attached base packages:
[1] grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] eNetXplorer_1.0.2    ComplexHeatmap_2.0.0 corrplot_0.84        fgsea_1.10.1         Rcpp_1.0.2          
 [6] BiocManager_1.30.4   circlize_0.4.8       readat_1.10.0        readxl_1.3.1         data.table_1.12.2   
[11] forcats_0.4.0        stringr_1.4.0        dplyr_0.8.3          purrr_0.3.2          readr_1.3.1         
[16] tidyr_0.8.3          tibble_2.1.3         ggplot2_3.2.0        tidyverse_1.2.1     

loaded via a namespace (and not attached):
 [1] nlme_3.1-141                bitops_1.0-6                matrixStats_0.54.0          lubridate_1.7.4            
 [5] progress_1.2.2              RColorBrewer_1.1-2          httr_1.4.1                  GenomeInfoDb_1.20.0        
 [9] tools_3.6.1                 backports_1.1.4             R6_2.4.0                    KernSmooth_2.23-15         
[13] lazyeval_0.2.2              BiocGenerics_0.30.0         colorspace_1.4-1            GetoptLong_0.1.7           
[17] withr_2.1.2                 assertive.reflection_0.0-4  prettyunits_1.0.2           tidyselect_0.2.5           
[21] gridExtra_2.3               compiler_3.6.1              glmnet_2.0-18               cli_1.1.0                  
[25] rvest_0.3.4                 Biobase_2.44.0              expm_0.999-4                assertive.properties_0.0-4 
[29] xml2_1.2.1                  DelayedArray_0.10.0         assertive.files_0.0-2       caTools_1.17.1.2           
[33] scales_1.0.0                XVector_0.24.0              assertive.numbers_0.0-2     pkgconfig_2.0.2            
[37] rlang_0.4.0                 GlobalOptions_0.1.0         rstudioapi_0.10             shape_1.4.4                
[41] generics_0.0.2              jsonlite_1.6                gtools_3.8.1                BiocParallel_1.17.18       
[45] zip_2.0.4                   RCurl_1.95-4.12             magrittr_1.5                GenomeInfoDbData_1.2.1     
[49] Matrix_1.2-17               munsell_0.5.0               S4Vectors_0.22.0            pathological_0.1-2         
[53] stringi_1.4.3               assertive.base_0.0-7        SummarizedExperiment_1.14.1 zlibbioc_1.30.0            
[57] gplots_3.0.1.1              plyr_1.8.4                  gdata_2.18.0                parallel_3.6.1             
[61] crayon_1.3.4                lattice_0.20-38             haven_2.1.1                 hms_0.5.1                  
[65] zeallot_0.1.0               pillar_1.4.2                GenomicRanges_1.36.0        rjson_0.2.20               
[69] assertive.sets_0.0-3        reshape2_1.4.3              codetools_0.2-16            stats4_3.6.1               
[73] fastmatch_1.1-0             glue_1.3.1                  calibrate_1.7.2             modelr_0.1.5               
[77] foreach_1.4.7               png_0.1-7                   vctrs_0.2.0                 testthat_2.2.1             
[81] cellranger_1.1.0            gtable_0.3.0                clue_0.3-57                 assertive.strings_0.0-3    
[85] assertthat_0.2.1            openxlsx_4.1.0.1            broom_0.5.2                 assertive.types_0.0-3      
[89] iterators_1.0.12            IRanges_2.18.2              cluster_2.1.0    

```
